---
title: "Campus Lagoon eBird data summary"
format: docx
editor: visual
author: Francis H. Joyce
---

Francis H. Joyce

13 February 2026

## Background

This is a quick analysis of eBird data available for the Campus Lagoon eBird hotspot ("[**UCSB--Campus Lagoon**](https://ebird.org/hotspot/L1072372)**" as of December 2025.**

The hotspot details page reports that **256 species have been recorded here (across all years), in 2,884 checklists, by 822 eBirders**.


Here I look at the number of species reported per year, the number of checklists submitted per year, and the relationship between those two variables.

Notes:

-   The code to generate this summary is in [this GitHub repository](https://github.com/ccber-restoration/sb-county-ebird). Because of the large file sizes, the repository does not contain the data, which can be downloaded from eBird after submitting a data download request.

-   This includes monthly CCBER surveys at Campus Lagoon if they were submitted to eBird

```{r}
#| message: false
#| warning: false
#| include: false

library(tidyverse)
library(cowplot)
#use ncos_data object 

cl_data <- read_rds("data-processed/cl_data.rds")

#note that infra-species taxa should be rolled up, this is just a rough cut
cl_species_richness <- cl_data %>% 
  #filter to category == species
  filter(category == "species") %>% 
  mutate(year = year(observation_date)) %>% 
  group_by(year) %>% 
  summarize(species_richness = n_distinct(common_name),
            n_checklists = n_distinct(checklist_id))

```


{{< pagebreak >}}

## Preliminary figures

```{r}
#| echo: false
#| label: fig-cl-sp_richness
#| fig-cap: Number of species recorded at the UCSB Campus Lagoon eBird hotspot per calendar year.
#plot over time
fig_cl_sr <- ggplot(data = cl_species_richness, aes(x = year, y = species_richness)) +
  geom_point() +
  theme_cowplot() +
  xlab("Year") +
  ylab("Species recorded")

fig_cl_sr

```

```{r}
#| echo: false
#| warning: false
#| label: fig-cl-sp_richness-recent
#| fig-cap: Number of species recorded per year between 2010 and 2025.
#plot over time
fig_cl_sr_recent <- ggplot(data = cl_species_richness, aes(x = year, y = species_richness)) +
  geom_point() +
  scale_x_continuous(limits = c(2010,2025), breaks = seq(2010, 2025, 2)) +
  theme_cowplot() +
  xlab("Year") +
  ylab("Species recorded")

fig_cl_sr_recent

```

```{r}
#| echo: false
#| label: fig-cl-checklists
#| fig-cap: Number of checklists submitted at the UCSB Campus Lagoon eBird hotspot per calendar year.

fig_cl_checklists <- ggplot(data = cl_species_richness, aes(x = year, y = n_checklists)) +
  geom_point() +
  theme_cowplot() +
  xlab("Year") +
  ylab("eBird Checklists")

fig_cl_checklists  

```

```{r}
#| echo: false
#| warning: false
#| label: fig-cl-checklists-recent
#| fig-cap: Number of checklists submitted at Campus Lagoon between 2010-2025. 

fig_cl_checklists_recent <- fig_cl_checklists <- ggplot(data = cl_species_richness, aes(x = year, y = n_checklists)) +
  geom_point() +
  scale_x_continuous(limits = c(2010, 2025), breaks = seq(2010, 2025, by = 2)) +
  theme_cowplot() +
  xlab("Year") +
  ylab("eBird Checklists")

fig_cl_checklists_recent  

```

```{r}
#| echo: false
#| label: fig-cl-sp-vs-checklists
#| fig-cap: Number of species recorded per year in relation to the number of checklists submitted per year.

fig_sp_vs_checklists_cl <- ggplot(data = cl_species_richness, aes(x = n_checklists, y = species_richness)) +
  geom_point() +
  theme_cowplot() +
  xlab("Checklists per year") +
  ylab("Species recorded per year")

fig_sp_vs_checklists_cl

```
