---
title: "NCOS ebird summary"
format: docx
editor: visual
---

Francis H. Joyce

30 January 2026

## Background

This is a quick analysis of eBird data available for the North Campus Open Space eBird hotspot ("[**UCSB North Campus Open Space (formerly Ocean Meadows Golf Course)**](https://ebird.org/hotspot/L820867)**" as of December 2025.**

The hotspot details page reports that **259 species have been recorded here, in 5,769 checklists, by 712 eBirders**.

Here I look at the number of species reported per year, the number of checklists submitted per year, and the relationship between those two variables.

Notes:

The code to generate this summary is in [this GitHub repository](https://github.com/ccber-restoration/sb-county-ebird). Because of the large file sizes, the repository does not contain the data, which can be downloaded from eBird after submitting a data download request.

```{r}
#| message: false
#| warning: false
#| include: false

library(tidyverse)
library(cowplot)
#use ncos_data object 

ncos_data <- read_rds("data-processed/ncos_data.rds")

#note that infra-species taxa should be rolled up, this is just a rough cut
ncos_species_richness <- ncos_data %>% 
  #filter to category == species
  filter(category == "species") %>% 
  mutate(year = year(observation_date)) %>% 
  group_by(year) %>% 
  summarize(species_richness = n_distinct(common_name),
            n_checklists = n_distinct(checklist_id))

```

{{< pagebreak >}}

## Preliminary figures

```{r}
#| echo: false

#plot over time
fig_ncos_sr <- ggplot(data = ncos_species_richness, aes(x = year, y = species_richness)) +
  geom_point() +
  geom_vline(xintercept = 2018, linetype = "dashed") +
  theme_cowplot() +
  xlab("Year") +
  ylab("Species recorded")
fig_ncos_sr

```

Figure 1. Number of species recorded at the NCOS eBird hotspot per calendar year. The vertical dashed line indicates 2018, when NCOS was opened to the public.

```{r}
#| echo: false
#| warning: false
#plot over time
fig_ncos_sr_recent <- ggplot(data = ncos_species_richness, aes(x = year, y = species_richness)) +
  geom_point() +
  geom_vline(xintercept = 2018, linetype = "dashed") +
  scale_x_continuous(limits = c(2010,2025), breaks = seq(2010, 2025, 2)) +
  theme_cowplot() +
  xlab("Year") +
  ylab("Species recorded")

fig_ncos_sr_recent

```

Figure 2. Number of species recorded per year between 2010 and 2025.

```{r}
#| echo: false

fig_ncos_checklists <- ggplot(data = ncos_species_richness, aes(x = year, y = n_checklists)) +
  geom_point() +
  geom_vline(xintercept = 2018, linetype = "dashed") +
  theme_cowplot() +
  xlab("Year") +
  ylab("eBird Checklists")

fig_ncos_checklists  

```

Figure 3. Number of checklists submitted at the NCOS eBird hotspot per calendar year. The vertical dashed line indicates 2018, when NCOS was opened to the public.

```{r}
#| echo: false
#| warning: false

fig_ncos_checklists_recent <- fig_ncos_checklists <- ggplot(data = ncos_species_richness, aes(x = year, y = n_checklists)) +
  geom_point() +
  geom_vline(xintercept = 2018, linetype = "dashed") +
  scale_x_continuous(limits = c(2010, 2025), breaks = seq(2010, 2025, by = 2)) +
  theme_cowplot() +
  xlab("Year") +
  ylab("eBird Checklists")

fig_ncos_checklists_recent  

```

Figure 4. Number of checklists submitted at NCOS between 2010-2025. The vertical dashed line indicates 2018, when NCOS was opened to the public.

```{r}
#| echo: false

fig_sp_vs_checklists <- ggplot(data = ncos_species_richness, aes(x = n_checklists, y = species_richness)) +
  geom_point() +
  theme_cowplot() +
  xlab("Checklists per year") +
  ylab("Species recorded per year")

fig_sp_vs_checklists

```

Figure 5. Number of species recorded per year in relation to the number of checklists submitted per year.
